R Notebook

### 2020-01-30 src/comb-maxent.Rmd
ii

  
```{r pre-install & load packages, include=FALSE}
options(java.parameters = "-Xmx2g" )#needs to run before dismo loaded - boosts memory for java to 1g
packages <- c("dismo", "stringr", "rJava", "raster", "ecospat", "bossMaps", "lattice", "tidyr", "enmSdm", "ggplot2", "plyr", "reshape2")
source("../src/ipak.R")
ipak(packages)
```

@@

@@

```{r individual maps}
#aea <- raster("../output/env/aea.tif") #the correct CRS
s <- shapefile("../data/env/landp/land-wgs84v2.shp") #made in arcgis
crs(s) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
#acrs <- aea@crs
mcp <- shapefile("../output/bio/minimum_convex_poly/mcp-wgs84.shp")
#ss <- raster::intersect(s, mcp)  #trim land file to mcp

plotlist <- list()

yrlst <- c("1998", "2006", "2013")
mthlst <- c("03", "06", "09", "12")
varlst <- c("temp", "salinity", "o2")

pn <- 1
for (y in 1:length(yrlst)){
  yr <- yrlst[y]
  fdr <- paste0("../data/env/ascii-original/ascii", yr, "/")
  for (m in 1:length(mthlst)){
    mt <- mthlst[m]
    for (v in 1:length(varlst)){
      var <- varlst[v]
      ras <- raster(paste0(fdr, yr, "_", mt, "_", var, ".nc_11.7737.asc"))
      lyrnm <- names(ras)
      crs(ras) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
      #ras <- projectRaster(ras, aea)
      ras <- raster::mask(ras, mcp)

      plotlist[[pn]] <- levelplot(ras, margin=FALSE, colorkey=list(space='bottom', axis.line=list(col='black')), par.settings=list(axis.line=list(col='transparent')), scales=list(draw=FALSE), col.regions=colorRampPalette(blue2red(20))) + latticeExtra::layer(sp.polygons(s, fill = "dark grey", col = "transparent") ) #changed ss to s
      pn <- pn+1

     # png(paste0("../output/maxent/paper_plots/", lyrnm, ".png"), height = 1024, width = 1024)
     # print(p)
     # dev.off()
    }
  }
}

#chl is done seperately as it is surface

yrlst <- c("1998", "2006", "2013")
mthlst <- c("03", "06", "09", "12")
varlst <- c("chl")

for (y in 1:length(yrlst)){
  yr <- yrlst[y]
  fdr <- paste0("../data/env/ascii-original/ascii", yr, "/")
  for (m in 1:length(mthlst)){
    mt <- mthlst[m]
    for (v in 1:length(varlst)){
      var <- varlst[v]
      ras <- raster(paste0(fdr, yr, "_", mt, "_", var, ".nc_0.50576.asc"))
      lyrnm <- names(ras)
      crs(ras) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
      ras <- raster::mask(ras, mcp)
      plotlist[[pn]] <- levelplot(ras, margin=FALSE, colorkey=list(space='bottom', axis.line=list(col='black')), par.settings=list(axis.line=list(col='transparent')), scales=list(draw=FALSE), col.regions=colorRampPalette(blue2red(20))) + latticeExtra::layer(sp.polygons(s, fill = "dark grey", col = "transparent"))
      pn <- pn+1

   #   png(paste0("../output/maxent/paper_plots/", lyrnm, ".png"), height = 1024, width = 1024)
    #  print(p)
    #  dev.off()
    }
  }
}
```

>>>>

  
```{r pre-install & load packages, include=FALSE}
options(java.parameters = "-Xmx2g" )#needs to run before dismo loaded - boosts memory for java to 1g
packages <- c("dismo", "stringr", "rJava", "raster", "ecospat", "bossMaps", "lattice", "tidyr", "enmSdm", "ggplot2", "plyr", "reshape2", "tidyverse", "patchwork", "purrr")
source("../src/ipak.R")
ipak(packages)
```

@@

# sample probabily maps

have moved sample files into a separate folder
march,  june, sep, dec
11.7737, 61.1128, 199.79. 628.026

```{r}
library(sf)
library(purrr)
library(raster)
library(lubridate)
library(tidyverse)

proj <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"

s <- st_read("../data/env/landp/land-wgs84v2.shp") %>% 
  st_set_crs(proj)

mcp <- st_read("../output/bio/minimum_convex_poly/mcp-wgs84.shp")

filelist <- list.files("../data/plot2-probmaps/",pattern="avg_cloglog*",recursive = TRUE,include.dirs = TRUE)

x <- stack(file.path("../data/plot2-probmaps/",filelist)) %>% 
  mask(mcp)

names(x) <- filelist

df <- x %>% 
  rasterToPoints() %>% 
  as.data.frame() %>% 
  gather(key = "variable", value = "value",-x,-y) %>% 
  mutate(variable=gsub(".asc","",variable),
         month=factor(months(parse_date_time(gsub("\\..*$", "",variable),"%m")),
                      ordered = TRUE,
                      levels=months(parse_date_time(1:12,"%m"))),
         depth=as.numeric(gsub("^.*_", "",variable)))

#remi code stops here

df$month <- str_sub(df$variable, start = 16, end = 18)

CapStr <- function(y) {
  c <- strsplit(y, " ")[[1]]
  paste(toupper(substring(c, 1,1)), substring(c, 2),
      sep="", collapse=" ")
}
df$month <- sapply(df$month, CapStr)

#remi code starts here

depthlab <- paste(unique(df$depth),"m")[c(1,3,2,4)]
names(depthlab) <- unique(df$depth)[c(1,3,2,4)]

ggplot(df)+
  geom_sf(data=mcp,fill='grey')+
  geom_raster(aes(x,y,fill=value))+
  geom_sf(data=mcp,fill="transparent")+
  scale_fill_viridis(option="magma",
                       direction=-1, limits = c(0,1), breaks = c(0, 0.25, 0.5, 0.75, 1),
                       guide=guide_colourbar(barwidth=0.5, barheight = 15, title = NULL, draw.ulim = FALSE, draw.llim = FALSE))+
  facet_grid(depth~month,
             labeller = labeller(depth=depthlab))+
  theme_classic()+
  scale_x_continuous(breaks = seq(-70,-50,10))+
  theme(legend.position = "right",
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.x=element_text(angle=45,vjust=1,hjust=1), strip.background = element_blank(),
        strip.text.x = element_text(face = "bold", size = 12),
        strip.text.y = element_text(face = "bold", size = 12))

ggsave("../output/maxent/paper_plots/test-overview.png", width = 210, height = 297, units = "mm", dpi = 600)
```

@@

### 2020-01-30 src/testing_file.Rmd
ii

df$month <- str_sub(df$variable, start = 16, end = 18)

@@

ggplot(df)+
  geom_sf(data=mcp,fill='grey')+
  geom_raster(aes(x,y,fill=value))+
  geom_sf(data=mcp,fill="transparent")+
  scale_fill_viridis(option="magma",
                       direction=-1,
                       guide=guide_colourbar(barwidth=20))+
  facet_grid(depth~month,
             labeller = labeller(depth=depthlab))+
  theme_classic()+
  scale_x_continuous(breaks = seq(-70,-50,10))+
  theme(legend.position = "bottom",
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.x=element_text(angle=45,vjust=1,hjust=1), strip.background = element_blank())
```

>>>>

#here i capitalise the months
df$month <- str_sub(df$variable, start = 16, end = 18)

@@

ggplot(df)+
  geom_sf(data=mcp,fill='grey')+
  geom_raster(aes(x,y,fill=value))+
  geom_sf(data=mcp,fill="transparent")+
  scale_fill_viridis(option="magma",
                       direction=-1, limits = c(0,1), breaks = c(0, 0.25, 0.5, 0.75, 1),
                       guide=guide_colourbar(barwidth=0.5, barheight = 15, title = NULL, draw.ulim = FALSE, draw.llim = FALSE))+
  facet_grid(depth~month,
             labeller = labeller(depth=depthlab))+
  theme_classic()+
  scale_x_continuous(breaks = seq(-70,-50,10))+
  theme(legend.position = "right",
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.x=element_text(angle=45,vjust=1,hjust=1), strip.background = element_blank(),
        strip.text.x = element_text(face = "bold", size = 12),
        strip.text.y = element_text(face = "bold", size = 12))

ggsave("../output/maxent/paper_plots/test-overview.png", width = 210, height = 297, units = "mm", dpi = 600)
```

#variable maps
```{r}

library(sf)
library(purrr)
library(raster)
library(lubridate)
library(patchwork)
library(tidyverse)

CapStr <- function(y) {
  c <- strsplit(y, " ")[[1]]
  paste(toupper(substring(c, 1,1)), substring(c, 2),
      sep="", collapse=" ")
}

mcp <- st_read("../data/plot4-variables/minimum_convex_poly/mcp-wgs84.shp")

s <- st_read("../data/plot4-variables/landp/land-wgs84v2.shp") %>% 
  st_set_crs(proj) %>% 
  st_intersection(mcp)

filelist <- list.files("../data/plot4-variables/ascii/ascii1998/",pattern="*.asc",recursive = TRUE)

x <- stack(file.path("../data/plot4-variables/ascii/ascii1998/",filelist)) %>% 
  mask(mcp) %>% 
  rasterToPoints() %>% 
  as.data.frame() %>% 
  gather(key = "variable", value = "value",-x,-y) %>% 
  separate(variable,c('year','month','variable','depth'),'_',remove=TRUE) %>% 
  mutate(month=factor(months(parse_date_time(month,"%m")),
                      ordered = TRUE,
                      levels=months(parse_date_time(1:12,"%m"))),
         year=as.numeric(gsub("X","",year)),
         monthyear=factor(paste0(month,'\n',year),
                          ordered = TRUE,
                          levels=unique(paste0(month,'\n',year))),
         variable=gsub(".nc","",variable)
         )

vars = c('chl','o2','salinity','temp')

for(v in vars){
  if(v=='chl'){
  assign(v,ggplot(x %>% filter(variable==v,month!='September'))+
           geom_sf(data=s,fill='grey')+
           geom_raster(aes(x,y,fill=value))+
           geom_sf(data=mcp,fill="transparent")+
           scale_fill_viridis(option="viridis", direction = -1 , 
           #scale_fill_distiller(palette="Spectral", direction=-1)+
           guide_colourbar(barwidth=0.5, title = "mmol\u207B\u00B3", draw.ulim = FALSE, draw.llim = FALSE))+

           facet_wrap(~monthyear,nrow=4,strip.position = "right")+
           theme_classic()+
           scale_x_continuous(breaks = seq(-70,-50,10))+
           labs(title=sapply(v, CapStr))+ #i added

           theme(legend.position = "bottom",
                 axis.title.x=element_blank(),
                 axis.title.y=element_blank(),
                 axis.text.x=element_text(angle=45,vjust=1,hjust=1),
                 strip.background = element_blank(),
                 strip.text = element_blank()))
  }else if(v=='o2'){
      assign(v,ggplot(x %>% filter(variable==v,month!='September'))+
           geom_sf(data=s,fill='grey')+
           geom_raster(aes(x,y,fill=value))+
           geom_sf(data=mcp,fill="transparent")+
           scale_fill_viridis(option="viridis", direction = -1 , 
           #scale_fill_distiller(palette="Spectral", direction=-1)+
           guide_colourbar(barwidth=0.5, title = "mmol\u207B\u00B3", draw.ulim = FALSE, draw.llim = FALSE))+
           labs(title=sapply(v, CapStr))+ #i added

           facet_wrap(~monthyear,nrow=4,strip.position = "right")+
           theme_classic()+
           scale_x_continuous(breaks = seq(-70,-50,10))+
           theme(legend.position = "bottom",
                 axis.title.x=element_blank(),
                 axis.title.y=element_blank(),
                 axis.text.x=element_text(angle=45,vjust=1,hjust=1),
                 strip.background = element_blank(),
                 strip.text = element_blank()))
  }else if(v=='salinity'){
      assign(v,ggplot(x %>% filter(variable==v,month!='September'))+
           geom_sf(data=s,fill='grey')+
           geom_raster(aes(x,y,fill=value))+
           geom_sf(data=mcp,fill="transparent")+
           scale_fill_viridis(option="viridis", direction = -1 , 
           #scale_fill_distiller(palette="Spectral", direction=-1)+
           guide_colourbar(barwidth=0.5, title = "PSU", draw.ulim = FALSE, draw.llim = FALSE))+

           facet_wrap(~monthyear,nrow=4,strip.position = "right")+
           labs(title=sapply(v, CapStr))+ #i added
           theme_classic()+
           scale_x_continuous(breaks = seq(-70,-50,10))+
           theme(legend.position = "bottom",
                 axis.title.x=element_blank(),
                 axis.title.y=element_blank(),
                 axis.text.x=element_text(angle=45,vjust=1,hjust=1),
                 strip.background = element_blank(),
                 strip.text = element_blank()))
  }
  else if(v=='temp'){
    assign(v,ggplot(x %>% filter(variable==v,month!='September'))+
             geom_sf(data=s,fill='grey')+
             geom_raster(aes(x,y,fill=value))+
             geom_sf(data=mcp,fill="transparent")+
             scale_fill_viridis(option="viridis", direction = -1, 
             #scale_fill_distiller(palette="Spectral", direction=-1)+
             guide_colourbar(barwidth=0.5, title = "Kelvin", draw.ulim = FALSE, draw.llim = FALSE))+
             facet_wrap(~monthyear,nrow=4,strip.position = "right")+
             labs(title=sapply(v, CapStr))+ #i added
             theme_classic()+
             scale_x_continuous(breaks = seq(-70,-50,10))+
             theme(legend.position = "bottom",
                   axis.title.x=element_blank(),
                   axis.title.y=element_blank(),
                   axis.text.x=element_text(angle=45,vjust=1,hjust=1), strip.background = element_blank(),
                   strip.text.y = element_text(face = "bold")))
  }
}

chl|o2|salinity|temp

ggsave("../output/maxent/paper_plots/variable_samples.png", height = 210, width = 297, units = "mm", dpi = 600)
```

```{r}
library(sf)
library(purrr)
library(raster)
library(lubridate)
library(patchwork)
library(tidyverse)

mcp <- st_read("../data/plot4-variables/minimum_convex_poly/mcp-wgs84.shp")

s <- st_read("../data/plot4-variables/landp/land-wgs84v2.shp") %>% 
  st_set_crs(proj) %>% 
  st_intersection(mcp)

filelist <- list.files("../data/plot4-variables/ascii/ascii1998/",pattern="*.asc",recursive = TRUE)

x <- stack(file.path("../data/plot4-variables/ascii/ascii1998/",filelist)) %>% 
  mask(mcp) %>% 
  rasterToPoints() %>% 
  as.data.frame() %>% 
  gather(key = "variable", value = "value",-x,-y) %>% 
  separate(variable,c('year','month','variable','depth'),'_',remove=TRUE) %>% 
  mutate(month=factor(months(parse_date_time(month,"%m")),
                      ordered = TRUE,
                      levels=months(parse_date_time(1:12,"%m"))),
         year=as.numeric(gsub("X","",year)),
         monthyear=factor(paste0(month,'\n',year),
                          ordered = TRUE,
                          levels=unique(paste0(month,'\n',year))),
         variable=gsub(".nc","",variable)
         )
vars = c('chl','o2','salinity','temp')
for(v in vars){
  assign(v,ggplot(x %>% filter(variable==v,month!='September'))+
           geom_sf(data=s,fill='grey')+
           geom_raster(aes(x,y,fill=value))+
           geom_sf(data=mcp,fill="transparent")+
scale_fill_viridis(option="viridis", direction = -1, 
             #scale_fill_distiller(palette="Spectral", direction=-1)+
             guide_colourbar(barwidth=0.5, title = "Kelvin", draw.ulim = FALSE, draw.llim = FALSE))+
           facet_wrap(~monthyear,nrow=4,strip.position = "right")+
           labs(title=v)+
           theme_classic()+
           scale_x_continuous(breaks = seq(-70,-50,10))+
           theme(legend.position = "bottom",
                 axis.title.x=element_blank(),
                 axis.title.y=element_blank(),
                 axis.text.x=element_text(angle=45,vjust=1,hjust=1),
                 strip.background = element_blank(),
                 strip.text = element_blank()))
  if(v=='temp'){
    assign(v,ggplot(x %>% filter(variable==v,month!='September'))+
             geom_sf(data=s,fill='grey')+
             geom_raster(aes(x,y,fill=value))+
             geom_sf(data=mcp,fill="transparent")+
scale_fill_viridis(option="viridis", direction = -1, 
             #scale_fill_distiller(palette="Spectral", direction=-1)+
             guide_colourbar(barwidth=0.5, title = "Kelvin", draw.ulim = FALSE, draw.llim = FALSE))+
             facet_wrap(~monthyear,nrow=4,strip.position = "right")+
             labs(title=v)+
             theme_classic()+
             scale_x_continuous(breaks = seq(-70,-50,10))+
             theme(legend.position = "bottom",
                   axis.title.x=element_blank(),
                   axis.title.y=element_blank(),
                   axis.text.x=element_text(angle=45,vjust=1,hjust=1)))
  }
}

chl|o2|salinity|temp
```

mcp <- st_read("../data/plot4-variables/minimum_convex_poly/mcp-wgs84.shp")

s <- st_read("../data/plot3-2d/landp/land-wgs84v2.shp") %>% 
  st_set_crs(proj) %>% 
  st_intersection(mcp)

filelist <- list.files("../data/plot3-2d/hotspot-mapping/",include.dirs = TRUE)

x <- stack(file.path("../data/plot3-2d/hotspot-mapping/",filelist)) %>% 
  mask(mcp) %>% 
  rasterToPoints() %>% 
  as.data.frame() %>% 
  gather(key = "variable", value = "value",-x,-y) %>% 
  mutate(month=factor(months(parse_date_time(gsub("_.*$", "",variable),"%m")),
                      ordered = TRUE,
                      levels=months(parse_date_time(1:12,"%m"))),
         value=ifelse(value==1,">0.5","<0.5"))

df <- x %>% 
  group_by(month) %>% 
  summarize(total=round(sum(value==">0.5")/n(),2))

ggplot(x)+
  geom_sf(data=s,fill='grey',colour='transparent')+
  geom_raster(aes(x,y,fill=value))+
  scale_fill_manual(values=c('transparent','red'))+
  geom_sf(data=mcp,fill="transparent")+
  geom_text(data=df,x=-50,y=67,label=df$total)+
  facet_wrap(~month,nrow=2)+#,
  theme_classic()+
  scale_x_continuous(breaks = seq(-70,-50,10))+
  theme(legend.position = "none",
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.x=element_text(angle=45,vjust=1,hjust=1))
```
